---
title: "House Prices Analysis and Prediction"
author: "Moaz M. El-Essawey"
date: "April 2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Packages and dataset
```{r, warning=FALSE}
library(tidyverse)

train_df <- read_csv("./data/train.csv")
head(train_df, 5)
```
```{r}
glimpse(train_df)
```

# Wranlge train Dataset
## Dropping NaN values
I will drop any column contains more than 200 NaN value in it.
and drop any non numeric column containing any NaN value.
and fill any numeric column with NaN values less than 200 with it's mean.
```{r}
nan_cols <- train_df %>% is.na() %>% colSums()
nan_cols <- nan_cols[nan_cols > 0]
nan_cols
```

```{r}
nan_cols_to_drop <- nan_cols[nan_cols >= 200]
train_df <- train_df %>% select(-names(nan_cols_to_drop))

nan_cols <- train_df %>% is.na() %>% colSums()
nan_cols <- nan_cols[nan_cols > 0]
nan_cols
```

```{r}
non_numeric_nan_cols <- colnames(train_df %>% select(names(nan_cols)) %>%  select(where(is.character)))
non_numeric_nan_cols


train_df <- train_df %>% select(-non_numeric_nan_cols)
train_df %>% head(5)
```
```{r}
nan_cols <- train_df %>% is.na() %>% colSums()
nan_cols[nan_cols > 0]
```

```{r}
train_df <- train_df %>% replace_na(
  list(MasVnrArea = mean(train_df$MasVnrArea, na.rm = TRUE),
       GarageYrBlt = mean(train_df$GarageYrBlt, na.rm = TRUE))
)

nan_cols <- train_df %>% is.na() %>% colSums()
nan_cols[nan_cols > 0]
```

Now our data frame does not contain any **NA** values. we can start working with and inspecting it's columns in details.

## Inspecting Categorical Columns
```{r}
categorical_cols <- colnames(train_df %>% select(where(is.character)))
numerical_cols <- colnames(train_df %>% select(where(is.numeric)))

print("Categorical Columns")
print(categorical_cols)
print("Numerical Columns")
print(numerical_cols)
```
### MSZoning Columns Analysis
```{r}
train_df %>% group_by(MSZoning) %>% summarise(counts=n())  %>% arrange(-counts)
```

**MSZoning** has a hig bias twards the **Residential Low Density** and **Residential Medium Density** zoning types.

### HouseStyle Column Analysis
```{r}
train_df %>% group_by(HouseStyle) %>% summarise(counts=n())  %>% arrange(-counts)
```

**HouseStyle** has major proportions twards **One Story, Two Story** and **One and one-half story**: *2nd level unfinished*

### SaleCondition Columns Analysis
```{r}
train_df %>% group_by(SaleCondition) %>% summarise(counts=n()) %>% arrange(-counts)
```

A House with **Normal** condition is dominated in selling condition.

## Cleaning Categorical Columns

```{r}
train_df <- train_df %>% mutate(
  low_density_zone = as.numeric(MSZoning == "RL"),
  
  one_story_type = as.numeric(HouseStyle == "1Story"),
  two_story_type = as.numeric(HouseStyle == "2Story"),
  half_story_type = as.numeric(HouseStyle == "1.5Fin"),
  
  normal_sale_cond = as.numeric(SaleCondition == "Normal"),
  gas_heating_sys = as.numeric(Heating == "GasA"),
  )

train_df %>% select(
  low_density_zone, one_story_type,two_story_type, half_story_type,
  normal_sale_cond, gas_heating_sys
) %>% head(5)
```

## Inscpecting Numerical Columns
```{r}
numerical_cols
```

### MSSubClass Column Analysis
```{r}
## convert MSSubClass to categorical column
train_df <- train_df %>% mutate(MSSubClass = as.character(MSSubClass))

train_df %>% group_by(MSSubClass) %>% 
  summarise(counts=n()) %>% arrange(-counts) %>% 
  ggplot(aes(x=MSSubClass, y=counts))+
    geom_col()+
    labs(
      x="MSSubClass",
      y="Count",
      title="Count of each Class in MSSubClass"
    )
```

most people prefer those types of houses
**20** :  **1-STORY 1946 & NEWER ALL STYLES**
**60** :  **2-STORY 1946 & NEWER**

### YearBuild, YearRemodedAdd and YrSold Columns Analysis
```{r}
train_df %>% mutate(
    time_taken_to_remodel = YearRemodAdd - YearBuilt,
    time_taken_to_sold = YrSold - YearBuilt,
    time_taken_to_sell_after_remodel = YrSold - YearRemodAdd) %>% 
  
  select(time_taken_to_remodel, time_taken_to_sold, time_taken_to_sell_after_remodel) %>% summary()
```

on average
 - it took around **13 years** to remodel a house.
 - it took around **36 years** to sell a house.
 - it took around **22 years** to sell a house that is remodeled.

### LotArea Column Analysis
```{r warning=FALSE}
print(train_df %>% select(LotArea) %>% summary)

train_df %>% ggplot(aes(x=LotArea))+
  geom_histogram(bins=40)+xlim(0, 50000)+
  labs(
    x="Lot Area",
    y="Count",
    title="Lot Area Dist."
  )
```

Most of the houses has a **Lot Area** around **5,000 to 15,000** square feet

### SalePrice Column Analysis
```{r}
print(train_df %>% select(SalePrice) %>% summary)

train_df %>% ggplot(aes(x=SalePrice))+
  geom_histogram(bins=40)+
  labs(
    x="Sale Price",
    y="Count",
    title="Sale Price Dist.")

train_df %>% ggplot(aes(x=SalePrice))+
  geom_boxplot()+
  labs(
    x="Sale Price",
    y="Count",
    title="Sale Price Dist.")
```

Most houses has price ranging between **1,000,000** to **2,500,000** USD.

## Correlation Matrix in the Dataset
```{r, fig.width=12, fig.height=11, fig.}
##install.packages("ggcorrplot")
library("ggcorrplot")
corr <- cor(train_df %>% select(where(is.numeric)))
corr[is.na(corr)] = 0

ggcorrplot(corr, hc.order = TRUE, outline.color = "white")
```

### SalePrice vs OverallCond
```{r}
train_df %>% mutate(OverallCond = as.character(OverallCond)) %>% 
  ggplot(aes(x=OverallCond, y=SalePrice, group=OverallCond))+
    geom_boxplot()+
    labs(
      x="Overall Condition",
      y="Sale Price",
      title="Sale Price by House Condition"
    )
```

### SalePrice vs TotalBsmtSF
```{r message=FALSE}
train_df %>% ggplot(aes(x=TotalBsmtSF, y=SalePrice))+
  geom_point(alpha=0.7)+
  geom_smooth(se=FALSE)
```

I detected an outliars at around more than **3,000** basement area, so i will choose to drop it.

```{r, warning=FALSE, message=FALSE}
train_df <- train_df %>% filter(TotalBsmtSF < 3000)

train_df %>% ggplot(aes(x=TotalBsmtSF, y=SalePrice))+
  geom_point(alpha=0.7)+
  geom_smooth(se=FALSE)+
  labs(
    x="Total Basement Area in SF",
    y="Sale Price",
    title="Sale Price by Basement Area"
  )

```

### SalePrice vs GrLivArea
```{r, message=FALSE}

train_df %>% ggplot(aes(x=GrLivArea, y=SalePrice))+
  geom_point(alpha=0.7)+
  geom_smooth(se=FALSE)+
  labs(
    x="Garden Area in SF",
    y="Sale Price",
    title="Sale Price by Garden Area"
  )

```

Both **TotalBsmtArea** and **GrLivArea** have very **strong positive** correlation with **SalePrice** columns.

# Machine Learning Models Analysis
**Ongoing....**